\mytitle

### 2016-11-22 pdocs/thesis/article-introduction.tex
article polishing
work on Aki's comments

% !TeX root = article.tex
\begin{abstract}
We introduce efficient method to simulate ductile fracture in existing physics engines.
\end{abstract}

@@

Computational methods used in physics engines are divided into modules that handle collision detection and 
contact description and modules that handle solution of equations of motion in real time. 
The equations of motion that need to be 
solved can be subdivided into equations related to motion, constraints and collisions. 
Velocity-based formulation is typically used in constraint based rigid body simulation, \cite{erleben.thesis}. 
Friction is typically taken into account and mechanical joints are handled by constraint equations,
\cite{erleben.thesis}.

Plasticity is not typically taken into account in gaming solutions and 
destruction of various bodies often takes place based on collision or impulse exceeding predefined limit.
Nevertheless, breaking of steel or reinforced concrete structures using this approach 
is not appropriate if the simulation is to look realistic. Theory for handling of plasticity 
has been presented already in \cite{cg1988}. \cite{muller2005meshless} 
present a method for modeling and animating of elastic and plastic bodies in real time using 
point based animation. This approach has not been widely used in computer games.  
One major issue is collision handling of deformable bodies.

This study will introduce an approach to account for plastic deformation in game applications.   
In the introduced method, plastic deformation takes place if the force or moment exceeds a predefined 
limit, deformation absorbs energy and joint breaks if plastic capacity is exceeded. 
The approach is based on using joint motors to model plasticity. 
The study extends a method introduced by
\cite{erleben.thesis} 
%\citet[p.~90]{erleben.thesis} 
which was originally proposed for modelling friction in joints. 
In the introduced method adjacent bodies are connected by motors. 
Motor power production limits are estimated based on the plastic section modulus. 
Joint breaking is based on summing plastic deformation and comparing it to a
predefined material based limit. The elastic part of deformation is modelled by employing 
a spring based on modification of an existing constraint in \cbullet.

>>>>

% !TeX root = article.tex
\begin{abstract}
This paper introduces simple and efficient method to simulate ductile fracture in existing physics engines.
Method is based on technique of splitting bodies to multiple pieces and joining them with constraints.
Constraints are described using body dimensions and material parameters.
Sample program with source code is made available to allow developers already using \cbullet\
to add plasticity into their simulations.
\end{abstract}

@@

Computational methods used in physics engines are divided into modules that handle collision detection and 
contact description and modules that handle solution of equations of motion in real time. 
The equations of motion that need to be 
solved can be subdivided into equations related to motion, constraints and collisions.

Velocity-based formulation is typically used in constraint based rigid body simulation, \cite{erleben.thesis}. 
Contacts and mechanical joints are handled by constraint equations 
and friction is typically taken into account,
\cite{erleben.thesis}.
Impulse-based rigid body simulation was introduced by \cite{Hahn:1998}, 
where time integrals of contact forces were used to model the interactions.
\cite{mirtich} suggested that all contact interactions
whether colliding, rolling, sliding or resting are affected
via trains of collisions.

Plasticity can be  accounted in gaming solutions by destructing various bodies based on collision
or impulse exceeding predefined limit.
Nevertheless, breaking of steel or reinforced concrete structures using this approach 
is not appropriate if the simulation is to look realistic. Theory for handling of plasticity 
has been presented already in \cite{cg1988}. 
\cite{Jones:2016:EPD} provides extensive listing of related work during past decades.
\cite{Obrien:2002:GMA} extends previously introduced finite element based method \cite{Obrien:1999:GMA}
to ductile materials.
\cite{muller2005meshless} 
present a method for modeling and animating of elastic and plastic bodies in real time using 
point based animation. This approach has not been widely used in computer games.  
One major issue is collision handling of deformable bodies.
Recently there has been multiple studies which have same target of bringing plasticity 
into wider usage.
\cite{Jones:2016:EPD} introduces an plasticity model which
maps impulses computed by rigid body solved to local example-based deformations.
\cite{Patkar:2015:EDB} presents a method for denting and bending of rigid bodies

\cite{Budsberg:2014:EPD}
Suggested methods are best suited for film production and high end games.

This study will introduce an approach to account for plastic deformation in game applications.   
Presented methodology does not require significant software development efforts from
game vendors and is thus easily adoptable. 
In the introduced method, plastic deformation takes place if the force or moment exceeds a predefined 
limit, deformation absorbs energy and joint breaks if plastic capacity is exceeded. 
The approach is based on using joint motors to model plasticity. 
The study extends a method introduced by
\cite{erleben.thesis} 
%\citet[p.~90]{erleben.thesis} 
which was originally proposed for modelling friction in joints. 
In the introduced method adjacent bodies are connected by motors. 
Motor power production limits are estimated based on the plastic section modulus. 
Joint breaking is based on summing plastic deformation and comparing it to a
predefined material based limit. The elastic part of deformation is modelled by employing 
a spring based on modification of an existing constraint in \cbullet.
